---
title: "My second Rmd document"
author: "Author: Myself"
date: "Last update: `r format(Sys.time(), '%d %b %Y')`"
output: 
    html_document:
        toc: true
        toc_depth: 2
        toc_float: true
        number_sections: true
        theme: default
        highlight: tango
---

```{r setup_chunk, echo=FALSE, message=FALSE}
	## load any packages you'll need in the document
	library(knitr)
	library(ggplot2)
	library(kableExtra)
	
	## here you can also load any custom-written functions you may need
	## source("~/Documents/R_scripts/my_custom_functions.R")
	
	## define global options for R code chunks
	opts_chunk$set(echo = FALSE, ## don't output the contents of the chunk
					message = FALSE, ## don't display R messages
					warning = FALSE, ## don't display R warnings
					fig.width=5, ## define standard figure widthd
					fig.height=5, ## define standard figure height
					fig.align = "center") ## define standard figure alignment
	## these can be changed in individual chunks if required (the options in
	## individual chunks will overwrite these), but if none are specified the 
	## chunks will default to these.
	
	## define colourblind palette for ggplot
	cbPalette <- c("#999999", "#CC79A7", "#b2182b", "#67001f", "#D55E00", "#E69F00", 
				   "#F0E442", "#7fbc41", "#009E73", "#56B4E9", "#0072B2", "#542788")

	## define custom ggplot theme
	custhm <- theme(axis.title.x = element_text(color="black", size=rel(1.5)),
					axis.text.x  = element_text(color="black", size=rel(1.5), 
												margin=margin(t = 10, b = 10)),
					axis.title.y = element_text(color="black", size=rel(1.5)),
					axis.text.y  = element_text(color="black", size=rel(1.5), 
												margin=margin(l = 10, r = 10)),
					legend.text = element_text(color="black", size=rel(1.2)),
					legend.position = "top",
					plot.title = element_text(size=rel(1.7), hjust=0.5, 
											  face="plain"),
					plot.subtitle = element_text(size=rel(1.1), 
												 hjust=0.5, face="plain"))
```

<!-- start of the document -->
<!-- comments in html are written like this :) -->

_This is a sample .Rmd file, which will investigate the built-in R dataset 
containing the average monthly temperatures in Nottingham (in Farenheights)._


# The dataset {.tabset .tabset-pills}

This is a random built-in dataset from R, I didn't even know it existed. 
It has average monthly temperatures between 1920 and 1939.

## A sneak peek into the dataset
```{r, echo=TRUE}
	typeof(nottem)
	nottem
```

## A better look using kable
```{r}
	nottemFrame <- cbind(c(1920:1939),
						 matrix(nottem, ncol=12, byrow=TRUE))

	auxheader <- 13
	names(auxheader) <- "Average monthly temperature"

	kable(nottemFrame, format="html", booktabs=TRUE, row.names=FALSE,
		  col.names=c(" ", "Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", 
		  			"Aug", "Sep", "Oct", "Nov", "Dec")) %>%
		kable_styling(full_width=FALSE, bootstrap_options=c("striped")) %>%
		column_spec(1, bold=TRUE, border_right=TRUE, color="blue") %>%
		row_spec(0, bold=TRUE, color="blue") %>%
		add_header_above(auxheader)
		
```

# Some plots {.tabset .tabset-pills}

```{r}
	## total number of years of observations
	nYears <- dim(nottemFrame)[1]
 	## prepare data.frame which will be used for plotting
	plotFrame <- matrix(0,12*nYears,3)
	## first column: nottemFrame as a single column vector, with removed $Years column
	## (this is equivalent to Jan 1920, Jan 1921, ..., Jan1939, Feb 1920, ..., Dec 1939
	## because I was too lazy to put it into chronological order - ggplot will be fine)
	plotFrame[,1] <- as.vector(nottemFrame[,-1])
	## second column: years 1920-1939
	plotFrame[,2] <- rep(1920:1939,12)
	## third column: months 1-12
	plotFrame[,3] <- as.vector(sapply(X=1:12,FUN=function(x){rep(x,nYears)}))
	## convert to data.frame() object, and name columns
	plotFrame <- data.frame(plotFrame)
	names(plotFrame) <- c("Temperature", "Year","Month")
		
	## create ggplot object
	myPlot <- ggplot(plotFrame) 
```

## A boxplot

```{r boxplot_example}
	myPlot +
		geom_boxplot(aes(x=as.factor(Month), y=Temperature)) +
		labs(x="Month", y="Average temperature (F)", title="Temperature in Nottingham")
```
This was trickier than I expected. Since the `plotFrame` dataset is all numeric,
if we simply call `geom_boxplot()` and ask for months on the x axis, it thinks 
that the scale is continuous. So it was necessary to convert `x=as.factor(Month)`
in the `aes()` plot call in order to get the desired results.


## A barplot

```{r barplot_example}
	myPlot +
		geom_col(aes(x=as.factor(Month), y=Temperature, fill=as.factor(Year))) +
		labs(x="Month", y="", title="Temperature in Nottingham") +
		guides(fill = guide_legend(title = "Year"))
```

## A histogram
```{r histogram_example}
	myPlot +
		geom_histogram(aes(x=Temperature, fill=as.factor(Month)), bins=25) +
		labs(x="Average monthly temperature", y="Frequency") +
		scale_fill_manual(values=cbPalette) +
		guides(fill = guide_legend(title = "Month"))
		
```