---
title: "Rmarkdown (and other things)"
author: rk595@cam.ac.uk
date: "Stats Meeting 26th Nov 2018"
fontsize: 15pt
output: 
    beamer_presentation:
        highlight: tango
        theme: "Boadilla"
        colortheme: "whale"
        fonttheme: "professionalfonts"
        slide_level: 2
        keep_tex: true
        toc: true
header-includes:
    \usepackage{booktabs}
    \definecolor{lightseagreen}{rgb}{0.15, 0.7, 0.65}
    \usecolortheme[named=lightseagreen]{structure}
    \hypersetup{colorlinks, urlcolor=lightseagreen, linkcolor=}
    \def\R{\includegraphics[width=0.04\textwidth]{./R_logo.png}}
    \def\begincols{\begin{columns}}
    \def\begincol{\begin{column}}
    \def\endcol{\end{column}}
    \def\endcols{\end{columns}}
---

```{r setup, echo=FALSE}
    library(ggplot2)
    library(kableExtra)
```

# Package 'ggplot'

## Package 'ggplot'

- Plot graphics in R

- More complicated than standard R graphics for simple plots...

- ... but easier for complex graphics / custom formatting.

- Plot formed of overlapping layers

- \alert{Requires a data.frame object}

## Example
\begin{figure}
	\includegraphics[height=0.8\textheight]{./irises.jpeg}
\end{figure}

## Example

### The Iris dataset
\footnotesize
```{r, echo=FALSE, eval=TRUE} 
head(iris, n=15)
```

## Example 
\footnotesize
\begincols
    \begincol{0.38\textwidth}
        \texttt{ggplot(iris)}
    \endcol
    \begincol{0.01\textwidth}
    \endcol
    \begincol{0.6\textwidth}
    ```{r, eval=TRUE, echo=FALSE}
        ggplot(iris)
    ``` 
    \endcol
\endcols

## Example 
\footnotesize
\begincols
    \begincol{0.38\textwidth}
        ```
        ggplot(iris) + geom_point(aes(x=Sepal.Length, y=Petal.Length)
        ```
    \endcol
    \begincol{0.01\textwidth}
    \endcol
    \begincol{0.6\textwidth}
    ```{r, eval=TRUE, echo=FALSE}
        ggplot(iris) +
        geom_point(aes(x=Sepal.Length, y=Petal.Length))
    ``` 
    \endcol
\endcols

## Example 
\footnotesize
\begincols
    \begincol{0.38\textwidth}
    ```
        ggplot(iris) +
        geom_point(aes(x=Sepal.Length, y=Petal.Length, col=Species), size=6) +
        geom_histogram(aes(x=Sepal.Length, y=..count.., alpha=0.5), bins = 20)
    ```
    \endcol
    \begincol{0.01\textwidth}
    \endcol
    \begincol{0.6\textwidth}
    ```{r, eval=TRUE, echo=FALSE}
        ggplot(iris) +
        geom_point(aes(x=Sepal.Length, y=Petal.Length, col=Species), size=6) +
        geom_histogram(aes(x=Sepal.Length, y=..count.., alpha=0.5), bins = 20)
    ``` 
    \endcol
\endcols

## Example 
\footnotesize
\begincols
    \begincol{0.38\textwidth}
    ```
    ggplot(iris) + geom_point(aes(x=Sepal.Length, y=Petal.Length, col=Species), size=6) + labs(x="Sepal length (cm)", y="Petal length (cm)", title="Iris flowers")
    ```
    \endcol
    \begincol{0.01\textwidth}
    \endcol
    \begincol{0.6\textwidth}
    ```{r, eval=TRUE, echo=FALSE}
        ggplot(iris) +
        geom_point(aes(x=Sepal.Length, y=Petal.Length, col=Species), size=6) +
        labs(x="Sepal length (cm)", y="Petal length (cm)", title="Iris flowers")
    ``` 
    \endcol
\endcols

## Example
\footnotesize
\begincols
    \begincol{0.38\textwidth}
        ```
        ggplot(iris) + geom_point(aes(x=Sepal.Length, y=Petal.Length, col=Species), size=6) +  labs(x="Sepal length (cm)", y="Petal length (cm)", title="Iris flowers") + theme(plot.title=element_text(size=50), axis.title.x=element_text(size=40, vjust=-1), axis.title.y=element_text(size=40, vjust=1), axis.text.x=element_text(size=30), axis.text.y=element_text(size=30), legend.title=element_blank(), legend.position = "top", legend.text=element_text(size=30), plot.margin = unit(c(1,1,1,1), "cm"))
        ```
    \endcol
    \begincol{0.01\textwidth}
    \endcol
    \begincol{0.6\textwidth}
    ```{r, eval=TRUE, echo=FALSE}
        ggplot(iris) +
        geom_point(aes(x=Sepal.Length, y=Petal.Length, col=Species), size=6) +
        labs(x="Sepal length (cm)", y="Petal length (cm)", title="Iris flowers") +
        theme(plot.title=element_text(size=50),
              axis.title.x=element_text(size=40, vjust=-1),
              axis.title.y=element_text(size=40, vjust=1),
              axis.text.x=element_text(size=30),
              axis.text.y=element_text(size=30),
              legend.title=element_blank(),
        	  legend.position = "top",
              legend.text=element_text(size=30),
              plot.margin = unit(c(1,1,1,1), "cm"))
    ``` 
    \endcol
\endcols

## Example

But it will be such a pain to write such a long code for each one of my plots...

\alert{Fear not!}

### {}
```
myTheme <- theme(plot.title=element_text(size=50),
              axis.title.x=element_text(size=40, vjust=-1),
              axis.title.y=element_text(size=40, vjust=1),
              axis.text.x=element_text(size=30),
              axis.text.y=element_text(size=30),
              legend.title=element_blank(),
           	  legend.position = "top",
              legend.text=element_text(size=30),
              plot.margin = unit(c(1,1,1,1), "cm"))
```

Then all you have to do is add a \alert{`+ myTheme`} at the end of your ggplot.

## Example
\footnotesize
\begincols
    \begincol{0.38\textwidth}
        ```
        ggplot(iris) +
        geom_point(aes(x=Sepal.Length, y=Petal.Length, col=Species), size=6) +
        labs(x="Sepal length (cm)", y="Petal length (cm)", title="Iris flowers") +
        myTheme
        ```
    \endcol
    \begincol{0.01\textwidth}
    \endcol
    \begincol{0.6\textwidth}
    ```{r, eval=TRUE, echo=FALSE}
        ggplot(iris) +
        geom_point(aes(x=Sepal.Length, y=Petal.Length, col=Species), size=6) +
        labs(x="Sepal length (cm)", y="Petal length (cm)", title="Iris flowers") +
        theme(plot.title=element_text(size=50),
              axis.title.x=element_text(size=40, vjust=-1),
              axis.title.y=element_text(size=40, vjust=1),
              axis.text.x=element_text(size=30),
              axis.text.y=element_text(size=30),
              legend.title=element_blank(),
          	  legend.position = "top",
              legend.text=element_text(size=30),
              plot.margin = unit(c(1,1,1,1), "cm"))
    ``` 
    \endcol
\endcols

## Resources

- \href{https://ggplot2.tidyverse.org/}{The Tidyverse official guide}

- \href{https://r4ds.had.co.nz/data-visualisation.html}{R for Data Science, chapter 3}

- \href{http://zevross.com/blog/2014/08/04/beautiful-plotting-in-r-a-ggplot2-cheatsheet-3/}{Detailed "cheat sheet" with many examples}

- \href{http://www.sthda.com/english/wiki/ggplot2-themes-and-background-colors-the-3-elements}{Themes and backgrounds}

- \href{http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/}{Colours (and colourblind-friendly palettes)}

- \href{http://www.cookbook-r.com/Graphs/Legends_(ggplot2)/}{Legends}

- \href{http://www.cookbook-r.com/Graphs/}{"Cookbook" for R - ggplot2}

- \href{https://www.rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf}{Summary cheat sheet}


<!-- ----------------------------------------------------------------------- -->

# Package 'kableExtra'

## Package 'kableExtra' 

- `kable` is a function from the `knitr` package
    + simple table generator

- `kableExtra` turns kable objects into fancier tables
    + exports to HTML and PDF
    + copy/paste from HTML often works to import tables to MS Word / Excel

- Allows for a "layer" syntax (sort of similar to ggplot)

## Example

First of all, we need a matrix with the data we want to output:

\tiny
```{r, echo=TRUE}
    sepalWidthF1 <- data.frame(table(iris$Sepal.Width[iris$Species=="setosa"]))
    sepalWidthF2 <- data.frame(table(iris$Sepal.Width[iris$Species=="versicolor"]))
    sepalWidthF3 <- data.frame(table(iris$Sepal.Width[iris$Species=="virginica"]))
    
    myTable <- merge(merge(sepalWidthF1, sepalWidthF2, by=1, all=TRUE),
          sepalWidthF3, by=1, all=TRUE)
    myTable <- myTable[order(as.numeric(as.character(myTable[,1]))),]
    myTable <- cbind(c("Sepal Width (cm)", "Setosa", "Versicolor", "Virginica"), t(myTable))
    
    myTable[is.na(myTable)] <- " "
    
    head(myTable)
```

## Example
```{r, echo=TRUE}
    myKable <- kable(myTable, format="latex", booktabs=TRUE, 
    				 row.names=FALSE, col.names=NULL)
    myKable
```

## Example
\footnotesize
```{r, echo=TRUE}
    theHeader <- c(1, 9, 10, 4)
    names(theHeader) <- c(" ", "Small", "Medium", "Large")
    
    myKable %>%
        kable_styling(full_width=FALSE, position="center", 
                      latex_options="scale_down") %>%
        group_rows("Species", 2, 4, latex_gap_space="1em") %>%
        add_header_above(theHeader) %>%
        footnote(general = "Here is a general comment about the table.")
```

## Resources

I'm sure there must be lots of excellent guides out there, but these are always my go-to places (and it often has the answers):

- \href{https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html}{Create awesome HTML tables with knitr:kable}

- \href{https://haozhu233.github.io/kableExtra/awesome_table_in_pdf.pdf}{Create awesome PDF tables with knitr:kable}

- \href{https://haozhu233.github.io/kableExtra/}{The kableExtra complete documentation}

<!-- ----------------------------------------------------------------------- -->

# Rmarkdown

## Rmarkdown

- Very useful for writing up drafts / reports / etc.

- Allows for code "chunks" to be run straight from \R 
    + If your data change, all of your figures and tables are automatically updated!
    
- Allows for flexible document structure
    + Tabs, navigation bar, etc.

- This presentation was written using the Rmarkdown/Beamer output!
    + It requires a bit of $\textrm{\LaTeX}$ knowledge...
    + ...but there are other interfaces available (e.g., _ioslides_, _slidy_, _xaringan_)

## Example: html document

### Header:
\footnotesize
```{r, eval=FALSE, echo=TRUE}
---
title: My first Rmd document
author: My name
date: "`r format(Sys.time(), '%d %b %Y')`" <- this displays current date
output: html_document:
    toc: true     <- four spaces - no tabs!
    toc_depth: 2
    number_sections: true
    theme: united
    highlight: tango
    ... 
---
```

## Example: html document

### Body:
\footnotesize
```{r, eval=FALSE, echo=TRUE}
After the header, you can just imput normal(ish) text.

- We can make lists by using hyphens
    + then four spaces to add subitems
    + like these
    
It's possible to make the font _italic_ or **bold** as well.

To change text colour inline, <span style="color:blue"> we need to use the HTML tag "span" </span>.

# It's easy to add section headers

## And subheaders as well {.tabset .tabset-pills}

### We can also make tabs

With something different in each tab

### I'm another tab

And this is my text

```

## Example: html document

### Body:
\footnotesize
```{r, eval=FALSE, echo=TRUE}
## But that's not all!

The really cool thing about Rmarkdown are the code chunks:
\```{r, echo=TRUE}
    x <- rnorm(100,0,1)
    y <- rnorm(100,0,1)
    plot(x,y)
\```

```

## Other formats / resources

- \href{https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf}{Rmarkdown cheat sheet}

- \href{https://bookdown.org/yihui/rmarkdown/word-document.html}{MS Word document}

- Rmarkdown also handles PDF outputs in a similar way, however, you probably need some knowledge of $\textrm{\LaTeX}$ to deal with the errors.

- \href{https://slides.yihui.name/xaringan/}{Xaringan - html presentations}

- \href{http://data-analytics.net/cep/Schedule_files/presentations_demo.html}{Slidy - html presentations}

- \href{https://bookdown.org/yihui/rmarkdown/ioslides-presentation.html}{Ioslides - html presentations}

- \href{https://www.dartistics.com/rmarkdown-shiny.html}{Shiny and Rmarkdown - responsive html}